### 2021-07-20 appendix/appendix.tex
background

\chapter{Box-plots of samples per project with all characteristics n*0.25 reduction}
\label{ap:full_25}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/full_25/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/full_25/boxplot_scribejava-core.png}}}
\end{figure}

@@

\chapter{Box-plots of samples per project with all characteristics n*0.50 reduction}
\label{ap:full_50}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/full_50/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/full_50/boxplot_scribejava-core.png}}}
\end{figure}

@@

\chapter{Box-plots of samples per project with all characteristics n*0.75 reduction}
\label{ap:full_75}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/full_75/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/full_75/boxplot_scribejava-core.png}}}
\end{figure}

@@

\chapter{Box-plots of samples per project without Levenshtein distance characteristic n*0.25 reduction}
\label{ap:no_distance_25}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_25/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_25/boxplot_scribejava-core.png}}}
\end{figure}

@@

\chapter{Box-plots of samples per project without Levenshtein distance characteristic n*0.50 reduction}
\label{ap:no_distance_50}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_50/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_50/boxplot_scribejava-core.png}}}
\end{figure}

@@

\chapter{Box-plots of samples per project without Levenshtein distance characteristic n*0.75 reduction}
\label{ap:no_distance_75}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_75/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_75/boxplot_scribejava-core.png}}}
\end{figure}

>>>>

\chapter{Box plots}
\section{Box-plots of samples per project with all characteristics with n*0.25 reduction}
\label{ap:full_25}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/full_25/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/full_25/boxplot_scribejava-core.png}}}
\end{figure}

@@

\section{Box-plots of samples per project with all characteristics with n*0.50 reduction}
\label{ap:full_50}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/full_50/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/full_50/boxplot_scribejava-core.png}}}
\end{figure}

@@

\section{Box-plots of samples per project with all characteristics with n*0.75 reduction}
\label{ap:full_75}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/full_75/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/full_75/boxplot_scribejava-core.png}}}
\end{figure}

@@

\section{Box-plots of samples per project without Levenshtein distance characteristic with n*0.25 reduction}
\label{ap:no_distance_25}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_25/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_25/boxplot_scribejava-core.png}}}
\end{figure}

@@

\section{Box-plots of samples per project without Levenshtein distance characteristic with n*0.50 reduction}
\label{ap:no_distance_50}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_50/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_50/boxplot_scribejava-core.png}}}
\end{figure}

@@

\section{Box-plots of samples per project without Levenshtein distance characteristic with n*0.75 reduction}
\label{ap:no_distance_75}
\begin{figure}[h]
    \centering
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_75/boxplot_google-auto-service.png}}}
    \qquad
    \subfloat{{\includegraphics[scale=0.32]{images/no_distance_75/boxplot_scribejava-core.png}}}
\end{figure}

### 2021-07-20 chapters/3_background.tex
background

\section{Clustering algorithms}
This research focuses on clustering mutants and use of existing clustering algorithms. 
Many different types of clustering algorithms have been proposed\cite{Rodriguez2019}. While there is a lot of diversity, some methods are more frequently used than others\cite{Wu2008TopMining}. 
Rodriguez et al., compared the performance of nine different clustering algorithms:
\begin{itemize}
    \item k-means
    \item CLARA
    \item hierarchical
    \item EM
    \item hcmodel
    \item spectral
    \item subscpace
    \item optics
    \item dbscan
\end{itemize}
The goal of their study was to guide researchers, who have little experience in data mining techniques, to the application of clustering methods.
They evaluated the algorithms in three distinct situations: default parameters, single parameter variation and random variation of parameters.
They used 400 generated artificial data sets which were normally distributed.
\newline
The results reported in their research are respective to specific configurations of normally distributed data and algorithmic implementations.
Nonetheless they do give a good overview on how the algorithms compare to each other.
\newline
For the default parameter experiment, the spectral clustering algorithm had the best performance and the hierarchical algorithm had the worst performance.
\newline
Regarding single parameter variations, for data sets containing 2 columns, the hierarchical, optics and EM methods showed significant performance variation.
\newline
With respect to the multidimensional analysis for data sets, the performance of the algorithms for the multidimensional selection of parameters was similar to that using the default parameters.
They conclude their research with observing that, for data sets with 10 or more columns the spectral algorithm consistently provided the best performance.
However the EM, hierarchical, k- means and subspace algorithms can also achieve similar performance with some parameter tuning.
The optics and dbscan algorithms aim at different data distributions than performed in this study. There different results could be obtained for non-normally distributed data.

>>>>

\section{Clustering algorithms}
This research focuses on clustering mutants and use of existing clustering algorithms. 
Many different types of clustering algorithms have been proposed\cite{Rodriguez2019}. While there is a lot of diversity, some methods are more frequently used than others\cite{Wu2008TopMining}. 
Rodriguez et al., compared the performance of nine different clustering algorithms:

\subsubsection{K-means} 
The K-means algorithm identifies \textit{k} number of centroids, and then allocates every data point to the nearest cluster, while keeping the centroids as small as possible\cite{Hartigan1979AlgorithmAlgorithm}.
\subsubsection{CLARA} 
CLARA uses multiple fixed samples of the data set to minimize sampling bias and, select the best medoids among those samples\cite{Sarle1991FindingAnalysis.}.
A medoid is defined as the object \textit{i} for which the average dissimilarity to all other objects in its cluster is minimal\cite{Sarle1991FindingAnalysis.}.
\subsubsection{Hierarchical clustering} 
Hierarchical clustering groups similar objects into groups called clusters\cite{Fraley1998HowAnalysis}. 
The endpoint is a set of clusters or groups, where each cluster is distinct from each other cluster, and the objects within each cluster are broadly similar to each other\cite{Fraley1998HowAnalysis}.
\subsubsection{EM} 
EM(expectation maximization) clustering technique is similar to the K-Means technique\cite{Fv1996TheAlgorithm}.
Instead of assigning examples to clusters to maximize the differences in means for continuous variables, the EM clustering algorithm computes probabilities of cluster memberships based on one or more probability distributions\cite{Fv1996TheAlgorithm}. 
\subsubsection{Spectral clustering}
Spectral clustering is a technique with roots in graph theory.
It identifies communities of nodes in a graph based on the edges connecting them\cite{Filippone2008AClustering}.
Spectral clustering uses information from the eigenvalues of special matrices built from the graph or the data set  to perform dimensionality reduction before clustering in fewer dimensions\cite{Filippone2008AClustering}. 
The similarity matrix is provided as an input and consists of a quantitative assessment of the relative similarity of each pair of points in the data set\cite{Filippone2008AClustering}.
\subsubsection{Subspace}
Subspace clustering algorithms consider the similarity between objects with respect to distinct subsets of the attributes\cite{Hartigan1979AlgorithmAlgorithm}. 
Different subsets of the attributes might define distinctions between each other. 
The algorithm can identify clusters that exist in multiple, possibly overlapping, sub spaces\cite{M2016SUBSPACEDATA}.
\subsubsection{Dbscan}
Dbscan is a density-based clustering non-parametric algorithm\cite{Daszykowski2009}.
Given a set of points in some space, it groups together points that are closely packed together, marking outliers as points that lie alone in low-density regions\cite{Daszykowski2009}.
\subsubsection{OPTICS}
The OPTICS(Ordering Points To Identify the Clustering Structure) algorithm starts with a data point and expands its neighborhood using a similar procedure as in the Dbscan algorithm\cite{Ankerst1999OPTICS:Structure}.
The difference is that the neighborhood is first expanded to points with low core-distance\cite{Ankerst1999OPTICS:Structure}. 
The optics algorithm can detect clusters having large density variations and irregular shapes\cite{Ankerst1999OPTICS:Structure}.
\newline

The goal of their study was to guide researchers, who have little experience in data mining techniques, to the application of clustering methods.
They evaluated the algorithms in three distinct situations: default parameters, single parameter variation and random variation of parameters.
They used 400 generated artificial data sets which were normally distributed.
\newline
The results reported in their research are respective to specific configurations of normally distributed data and algorithmic implementations.
Nonetheless they do give a good overview on how the algorithms compare to each other.
\newline
For the default parameter experiment, the spectral clustering algorithm had the best performance and the hierarchical algorithm had the worst performance.
\newline
Regarding single parameter variations, for data sets containing 2 columns, the hierarchical, optics and EM methods showed significant performance variation.
\newline
With respect to the multidimensional analysis for data sets, the performance of the algorithms for the multidimensional selection of parameters was similar to that using the default parameters.
They conclude their research with observing that, for data sets with 10 or more columns the spectral algorithm consistently provided the best performance.
However the EM, hierarchical, k- means and subspace algorithms can also achieve similar performance with some parameter tuning.
The optics and dbscan algorithms aim at different data distributions than performed in this study. There different results could be obtained for non-normally distributed data.

### 2021-07-20 chapters/rq1/2_clustering_mutants.tex
minor changes
background

\subsection{Hierarchical clustering}
Clustering Levenshtein distances has been done before with hierarchical clustering\cite{Rajalingam2011, Gothai2010PerformanceAlgorithms}. 
Research shows that hierarchical clustering performs better when clustering with at least ten features\cite{Rodriguez2019}.
It also states that varying the parameters of hierarchical clustering improves the performance compared to that of the default settings of the algorithm\cite{Rodriguez2019}.
We identified more than ten characteristics and can use them as features.
We adjust the parameters of the algorithm based on the characteristics in our data set.
Bases on these observations we select hierarchical clustering to cluster the mutants.
Next we will explain the configuration we use for clustering the mutants. 
\newline
Hierarchical clustering is subdivided into agglomerative and divisive. 
The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
Hierarchical clustering uses different metrics which calculates the euclidean distance between two clusters and the linkage criteria\cite{Rajalingam2011}. 
The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011}.
We reviewed the different linkage criteria and concluded the following.
\newline
Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
\newline
We identify all characteristics per mutant as a separate cluster.
Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
In summary, the agglomerative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.

@@

\subsection{Cluster size}
Agglomerative hierarchical clustering will continue to cluster until there is one cluster left. 
Naturally this cluster will contain all the mutants.
We can cut off the clustering algorithm at any point.
We decide on the number of clusters based on the amount of mutants generated in a full set.
The reasoning for the selecting the number of clusters will be from a performance reducing perspective.
Mutation clustering increases the performance by reducing the amount of mutants executed\cite{Pizzoleto2019}.
Reducing the performance by a quarter, half or three quarters will give an indication of performance increases.
To translate this to cluster size we will reduce the number of clusters in steps op 25\%.
In other words by reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014}.
If necessary we can decide to cut off the clustering at different points.
The baseline will consist of three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
The amount of mutants inside a cluster will be decided by the clustering algorithm.
\end{document}

>>>>

\subsection{Hierarchical clustering}
Clustering Levenshtein distances has been done before with hierarchical clustering with promising results.\cite{Rajalingam2011, Gothai2010PerformanceAlgorithms}. 
Research shows that hierarchical clustering performs better when clustering with at least ten features\cite{Rodriguez2019}.
It also states that varying the parameters of hierarchical clustering improves the performance compared to that of the default settings of the algorithm\cite{Rodriguez2019}.
We identified more than ten characteristics and can use them as features.
We adjust the parameters of the algorithm based on the characteristics in our data set.
Bases on these observations and the research of Rodriguez et al., we select hierarchical clustering as it gives us the best performance in terms of partitioning.
Next we will explain the configuration we use for clustering the mutants. 
\newline
Hierarchical clustering is subdivided into agglomerative and divisive. 
The agglomerative hierarchical technique follows bottom up approach whereas divisive follows top-down approaches.
Hierarchical clustering uses different metrics which calculates the euclidean distance between two clusters and the linkage criteria\cite{Rajalingam2011}. 
The linkage criteria specifies the dissimilarity in the sets as a function of the pair-wise distances of observations in those sets\cite{Rajalingam2011}.
We reviewed the different linkage criteria and concluded the following.
\newline
Research shows that the complete linkage outperforms the single linkage method\cite{Vijaya2019ComparativeClustering}.
The ward linkage and complete linkage methods perform the same when clusters are well separated\cite{Vijaya2019ComparativeClustering}.
However if the clusters overlap the ward linkage outperforms the complete linkage\cite{Vijaya2019ComparativeClustering}.
\newline
We identify all characteristics per mutant as a separate cluster.
Starting out with each mutant as a separate cluster we can use the agglomerative form of hierarchical clustering.
Since we cannot assume that our clusters are well separated we chose to use the ward linkage method for our clustering algorithm.
In summary, the agglomerative hierarchical clustering algorithm with ward linkage will cluster our mutants represented by the characteristics we gathered.

@@

\subsection{Cluster size}
Agglomerative hierarchical clustering will continue to cluster until there is one cluster left. 
Naturally this cluster will contain all the mutants.
We can cut off the clustering algorithm at any point.
We decide on the number of clusters based on the amount of mutants generated in a full set.
The reasoning for the selecting the number of clusters will be from a performance reducing perspective.
Mutation clustering increases the performance by reducing the amount of mutants executed\cite{Pizzoleto2019}.
Reducing the performance by a quarter, half or three quarters will give an indication of performance increases.
To translate this to cluster size we will reduce the number of clusters in steps op 25\%.
In other words by reducing the amount of mutants, for example by half, we increase the performance by 50\%\cite{Just2014AreTesting}.
If necessary we can decide to cut off the clustering at different points.
The baseline will consist of three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
The amount of mutants inside a cluster will be decided by the clustering algorithm.
\end{document}
